---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rrd

The `rrd` package allows you to read data from an [RRD](http://oss.oetiker.ch/rrdtool/) database.

Internally it uses [librrd](http://oss.oetiker.ch/rrdtool/doc/librrd.en.html) to import the binary data directly into R without exporting it to an intermediate format first.

For an introduction to RRD database, see https://oss.oetiker.ch/rrdtool/tut/rrd-beginners.en.html

## Installation


### Pre-requisites

In order to build the package from source you need [librrd](http://oss.oetiker.ch/rrdtool/doc/librrd.en.html).
Installing [RRDtool](http://oss.oetiker.ch/rrdtool/) from your package manager will usually also install the library.

In ubuntu:

```sh
sudo apt-get install rrdtool librrd-dev
```

In RHEL / CentOS:

```sh
sudo yum install rrdtool rrdtool-devel
```

### Installing from CRAN

`rrd` is not yet on CRAN

### Installing from github

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("andrie/rrd")
```

## Example


In R:

```{r load}
library(rrd)
```

To describe the contents of an RRD file, use `describe_rrd()`:


```{r describe}
rrd_cpu_0 <- system.file("extdata/cpu-0.rrd", package = "rrd")

describe_rrd(rrd_cpu_0)
```


To read an entire RRD file, i.e. all of the RRA archives, use `read_rrd()`. This returns a list of `tibble` objects:


```{r rrd}
cpu <- read_rrd(rrd_cpu_0)

str(cpu, max.level = 1)
```


Since the resulting object is a list of `tibble`s, you can easily work with individual data frames:

```{r inspect-rrd}
names(cpu)

cpu[[1]]

tail(cpu$AVERAGE60$sys)
```


To read a single RRA archive from an RRD file, use `read_rra()`. To use this function, you must specify several arguments that define the specific data to retrieve. This includes the consolidation function (e.g. "AVERAGE") and time step (e.g. 60), the `end` time.  You must also specifiy either the `start` time, or the number of steps, `n_steps`.

In this example, you extract the average for 1 minute periods (`step = 60`), for one entire day (`n_steps = 24 * 60`):

```{r rra}
end_time <- as.POSIXct("2018-05-02") # timestamp with data in example
avg_60 <- read_rra(rrd_cpu_0, cf = "AVERAGE", step = 60, n_steps = 24 * 60,
                     end = end_time)

avg_60
```

And you can easily plot using your favourite packages:

```{r, fig.height=3}
library(ggplot2)
ggplot(avg_60, aes(x = timestamp, y = user)) + 
  geom_line() +
  stat_smooth(method = "loess", span = 0.125, se = FALSE) +
  ggtitle("CPU0 usage, data read from RRD file")
```




## More information

For more information on `rrdtool` and the `rrd` format please refer to the official [rrdtool documentation](http://oss.oetiker.ch/rrdtool/doc/index.en.html) and [tutorials](http://oss.oetiker.ch/rrdtool/tut/index.en.html).

You can read a more in-depth description of the package and more examples in this [blog post](http://plamendimitrov.net/blog/2014/08/09/r-package-for-working-with-rrd-files/).
