---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Rrd

The `Rrd` package allows you to read data from an [RRD](http://oss.oetiker.ch/rrdtool/) database.

Internally it uses [librrd](http://oss.oetiker.ch/rrdtool/doc/librrd.en.html) to import the binary data directly into R without exporting it to an intermediate format first.

For an introduction to RRD database, see https://oss.oetiker.ch/rrdtool/tut/rrd-beginners.en.html

## Installation


## Pre-requisites

In order to build the package from source you need [librrd](http://oss.oetiker.ch/rrdtool/doc/librrd.en.html).
Installing [RRDtool](http://oss.oetiker.ch/rrdtool/) from your package manager will usually also install the library.

In ubuntu:

```sh
sudo apt-get install librrd-dev
```

In RHEL / CentOD:

```sh
sudo yum install librrd-dev
```


`Rrd` is not yet on CRAN

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("andrie/Rrd")
```

## Example


In R:

```{r load}
library(Rrd)
```

### Read an entire RRD file:

Use `read_rrd()`:


```{r rrd}
rrd_cpu_0 <- system.file("extdata/cpu-0.rrd", package = "Rrd")

cpu <- read_rrd(rrd_cpu_0)
names(cpu)
head(cpu[[1]])
tail(cpu[[1]])

tail(cpu$AVERAGE60$sys)
```


### Read a single RRA archive from an RRD file

Use `read_rra()`:

```{r rra}
rrd_cpu_0 <- system.file("extdata/cpu-0.rrd", package = "Rrd")

start_time <- as.POSIXct("2018-05-01") # timestamp with data in example
end_time <- as.POSIXct("2018-05-02") # timestamp with data in example
avg_60 <- read_rra(rrd_cpu_0, cf = "AVERAGE", step = 60L, 
                     start = start_time, 
                     end = end_time)
names(avg_60)
head(avg_60)
tail(avg_60)
```


## More information

For more information on `rrdtool` and the `rrd` format please refer to the official [rrdtool documentation](http://oss.oetiker.ch/rrdtool/doc/index.en.html) and [tutorials](http://oss.oetiker.ch/rrdtool/tut/index.en.html).

You can read a more in-depth description of the package and more examples in this [blog post](http://plamendimitrov.net/blog/2014/08/09/r-package-for-working-with-rrd-files/).
